---
title: "EDUC | 19S | Midterm"
author: "Teresa Chen"
date: "4/26/2019"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE,
                      wwarning = TRUE,
                      cache = TRUE)
```


## Part A: Data

```{r}
download_file <- function(year) {
    link <- glue::glue("https://www.oregon.gov/ode/educator-resources/assessment/TestResults20{year}/pagr_schools_ela_raceethnicity_{year-1}{year}.xlsx")
    rio::import(link, setclass = "tibble", na = c("-", "--", "*"))
}
```

### 1. Data preparation

Use the function above to download all the data for each of the past 4 school years and bind it into a single data frame, using a single function (i.e., one line of code). Note, this may take a minute or two to run, depending on your Internet speed.

```{r}

original <- map_df(c(15:18), download_file) 


```


Conduct some basic data cleaning to make your data file look like the following.

+ Filter for only student groups coded as "White" or "Hispanic/Latino".  
+ Select variables related to the number of students in each of the levels (1:4), and not percentages or collapsed levels.  
+ Remove any row that has missing data in any of the n variables

```{r}
df <- janitor::clean_names(original) %>% 
  filter(student_group == "White" | student_group == "Hispanic/Latino") %>% 
  dplyr::select(academic_year, district, school, student_group, grade_level, 
         number_level_4, number_level_3, number_level_2, number_level_1) %>% 
  gather(level, n, number_level_4:number_level_1) %>% 
  filter(!is.na(n)) %>% 
  mutate(level = parse_number(level)) %>% 
  arrange(academic_year, district, student_group, grade_level, level)

df
```

### 2. Cumulative calculation

Calculate the cumulative n for each school by student group, grade, and academic year. The result should look like the below. Hint, look at `?base::cumsum`.

```{r}
df_cn <- df %>% group_by(academic_year, school, student_group, grade_level) %>% 
  mutate(cn = cumsum(n))

df_cn
```

### 3. Data reformation

Reformat the data so it looks like the below, removing n and filling by cn. Remove rows that have missing data for either student group.

```{r}
# df_cn %>% 
  
```


# Part B






